A Hexagon Tile Map Algorithm for Displaying Spatial Data

### 2020-11-13 kobakian-cook.Rmd
figure captions

```{r choro, fig.cap = "A choropleth map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of Thyroid cancer for females, ranging from much lower than average (blue) to much higher than average (red)"}
aus_ggchoro <- ggplot(sa2thyroid_ERP) + 
  geom_sf(aes(fill = SIR), size = 0.1) + 
  scale_fill_identity() + invthm
aus_ggchoro
```

@@

```{r fullhexmap, fig.cap = "A hexagon tile map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of thyroid cancer for females, ranging from much lower than average (blue) to much higher than average (red)"}
if (!file.exists("data/aus_hexmap.rda")) {

@@

```{r tas_displays, width = 10, height = 3, fig.cap = "The three displays show alternative maps of the Statistical Areas at Level 2 in the Australian state of Tasmania. These alternatives show the (a) contiguous cartogram, (b) non-contiguous cartogram and (c) Dorling cartogram of Tasmania."}
library(gridExtra)
grid.arrange(tas_ggcont, tas_ggncont, tas_ggdorl, ncol=3)
```

@@

```{r hexmap, fig.cap = "A hexagon tile map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of thyroid for females, they range from much lower than average (blue) to much higher than average (red)."}
aus_hexmap_plot
```

@@

\begin{figure}
\includegraphics[width=14cm]{figs/sugarbag flow.png}
\caption{\label{fig:sugarbag_flow}A flow diagram detailing the necessary steps to create a hexagon tile map.}
\end{figure}

@@

```{r filterprocess, fig.cap = "Filter for grid points within a square, then circular, distance for those closest to the centroid.", dpi = 600, echo = FALSE, height = 12}

>>>>

```{r choro, fig.cap = "A choropleth map of thyroid incidence among females across the Statistical Areas of Australia at Level 2. Blue indicates lower than average and red indicates higher than average incidence. A cluster of high incidence is visible on the east coast."}
aus_ggchoro <- ggplot(sa2thyroid_ERP) + 
  geom_sf(aes(fill = SIR), size = 0.1) + 
  scale_fill_identity() + invthm
aus_ggchoro
```

@@

```{r fullhexmap, fig.cap = "A hexagon tile map of Australia at SA2 level. The colours communicate the value of the estimated SIR of thyroid cancer for females, ranging from much lower than average (blue) to much higher than average (red)"}
if (!file.exists("data/aus_hexmap.rda")) {

@@

```{r tas_displays, width = 10, height = 3, fig.cap = "The three displays show alternative maps of the Australian state of Tasmania at SA2 level: (a) contiguous cartogram, (b) non-contiguous cartogram and (c) Dorling cartogram of Tasmania. The contiguous cartogram looks like the state has an hourglass figure, while the non-contiguous cartogram shrinks areas into invisibility. The Dorling expands the metropolitan regions."}
library(gridExtra)
grid.arrange(tas_ggcont, tas_ggncont, tas_ggdorl, ncol=3)
```

@@

```{r hexmap, fig.cap = "A hexagon tile map of female tyroid cancer incidence in Australia, the same data as shown in the choropleth map in Figure 1. The high incidence in several of the metropolitan regions (Brisbane, Sydney and Perth) can now be seen, along with numerous isolated spots."}
aus_hexmap_plot
```

@@

\begin{figure}
\includegraphics[width=14cm]{figs/sugarbag flow.png}
\caption{\label{fig:sugarbag_flow}A flow diagram detailing the steps taken to create a hexagon tile map. There are two basic processes, one to make the grid, and the other to allocate centroids to grid points.}
\end{figure}

@@

```{r filterprocess, fig.cap = "Illustration of several steps of the algorithm: (a) ... (b) ... (c) ... (d). ", dpi = 600, echo = FALSE, height = 12}

### 2020-11-13 kobakian-cook.tex
figure captions

\begin{Schunk}
\begin{figure}
\includegraphics[width=0.95\linewidth]{kobakian-cook_files/figure-latex/choro-1} \caption[A choropleth map of the Statistical Areas of Australia at Level 2]{A choropleth map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of Thyroid cancer for females, ranging from much lower than average (blue) to much higher than average (red)}\label{fig:choro}
\end{figure}
\end{Schunk}

@@

\begin{Schunk}
\begin{figure}
\includegraphics[width=0.95\linewidth]{kobakian-cook_files/figure-latex/tas_displays-1} \caption[The three displays show alternative maps of the Statistical Areas at Level 2 in the Australian state of Tasmania]{The three displays show alternative maps of the Statistical Areas at Level 2 in the Australian state of Tasmania. These alternatives show the (a) contiguous cartogram, (b) non-contiguous cartogram and (c) Dorling cartogram of Tasmania.}\label{fig:tas_displays}
\end{figure}
\end{Schunk}

@@

\begin{Schunk}
\begin{figure}
\includegraphics[width=0.95\linewidth]{kobakian-cook_files/figure-latex/hexmap-1} \caption[A hexagon tile map of the Statistical Areas of Australia at Level 2]{A hexagon tile map of the Statistical Areas of Australia at Level 2. The colours communicate the value of the estimated SIR of thyroid for females, they range from much lower than average (blue) to much higher than average (red).}\label{fig:hexmap}
\end{figure}
\end{Schunk}

@@

\begin{Schunk}
\begin{figure}
\includegraphics[width=0.95\linewidth]{kobakian-cook_files/figure-latex/filterprocess-1} \caption[Filter for grid points within a square, then circular, distance for those closest to the centroid]{Filter for grid points within a square, then circular, distance for those closest to the centroid.}\label{fig:filterprocess}
\end{figure}
\end{Schunk}

@@

A choropleth map display is the current common practice for spatial data
visualisation. As the choropleth map display relies on the geographic
shapes, it does not effectively present the density of the population in
small geographic areas. The common solution to enhance the communication
of spatial distributions is to incorporate the population values as a
cartogram or another alternative map display. Many alternative map
displays do not work well for Australia due to the large rural
geographic space between the densely populated capital cities. The
administrative boundaries may also distract from the statistics
communicated using colour due to the small size of some areas. This
paper provides an algorithm for Australia called the hexagon tile map.

@@

This algorithm should be general enough to apply to other countries and
problems

This algorithm has only been tested using single countries, and does not
consider definite borders of countries. While the buffer allows
extension beyond the furthest centroids, there is no mechanism to
protect the borders and ensure centroids are placed within the
geographic borders of a country.

>>>>

\begin{Schunk}
\begin{figure}
\includegraphics[width=0.95\linewidth]{kobakian-cook_files/figure-latex/choro-1} \caption[A choropleth map of thyroid incidence among females across the Statistical Areas of Australia at Level 2]{A choropleth map of thyroid incidence among females across the Statistical Areas of Australia at Level 2. Blue indicates lower than average and red indicates higher than average incidence. A cluster of high incidence is visible on the east coast.}\label{fig:choro}
\end{figure}
\end{Schunk}

@@

\begin{Schunk}
\begin{figure}
\includegraphics[width=0.95\linewidth]{kobakian-cook_files/figure-latex/tas_displays-1} \caption[The three displays show alternative maps of the Australian state of Tasmania at SA2 level]{The three displays show alternative maps of the Australian state of Tasmania at SA2 level: (a) contiguous cartogram, (b) non-contiguous cartogram and (c) Dorling cartogram of Tasmania. The contiguous cartogram looks like the state has an hourglass figure, while the non-contiguous cartogram shrinks areas into invisibility. The Dorling expands the metropolitan regions.}\label{fig:tas_displays}
\end{figure}
\end{Schunk}

@@

\begin{Schunk}
\begin{figure}
\includegraphics[width=0.95\linewidth]{kobakian-cook_files/figure-latex/hexmap-1} \caption[A hexagon tile map of female tyroid cancer incidence in Australia, the same data as shown in the choropleth map in Figure 1]{A hexagon tile map of female tyroid cancer incidence in Australia, the same data as shown in the choropleth map in Figure 1. The high incidence in several of the metropolitan regions (Brisbane, Sydney and Perth) can now be seen, along with numerous isolated spots.}\label{fig:hexmap}
\end{figure}
\end{Schunk}

@@

\begin{Schunk}
\begin{figure}
\includegraphics[width=0.95\linewidth]{kobakian-cook_files/figure-latex/filterprocess-1} \caption[Illustration of several steps of the algorithm]{Illustration of several steps of the algorithm: (a) ... (b) ... (c) ... (d). }\label{fig:filterprocess}
\end{figure}
\end{Schunk}

@@

This paper has described a new algorithm for creating an alternative
display. This paper provides an algorithm for Australia called the
hexagon tile map.

A choropleth map display is the current common practice for spatial data
visualisation. As the choropleth map display relies on the geographic
shapes, it does not effectively present the density of the population in
small geographic areas. The common solution to enhance the communication
of spatial distributions is to incorporate the population values as a
cartogram or another alternative map display. Many alternative map
displays do not work well for Australia due to the large rural
geographic space between the densely populated capital cities. The
administrative boundaries may also distract from the statistics
communicated using colour due to the small size of some areas.

@@

This algorithm should be general enough to apply to other countries and
problems. XXX This algorithm has only been tested using single
countries, and does not consider definite borders of countries. While
the buffer allows extension beyond the furthest centroids, there is no
mechanism to protect the borders and ensure centroids are placed within
the geographic borders of a country.
